---
title: "covid 19"
author: "danitza Nunez"
date: "10/12/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex


$H_0$: Dying is not independent of sex.

$H_A$ : Men and women die at different rates, dying is dependent on sex. 


Based on the variables, the helper file being used is CAT~CAT because both the response variable and explanatory variables are categorical. 

Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

The results provided show 25 females died out of the 1831. Among the males, 44 died out of the 1411. Based on the data, males have a higher death rate. 

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

`
```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```
  
Among the individuals that died of COVID, the plot shows that males are most likely to die than females. 




  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

Twice the amount of males died compared to females. 


## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The Chi-squared test shows the P-Value = 0.0012 which is less than 0.05 which means we reject the null hypothesis. In the Fisher Test the p-value is 0.0009 which means the odds are that males in Korea are 2.3 times more likely to die than females. 



## Conclusions

Men and women die at different rates, hence we rejected the null hypothesis  because in the Fisher test the p-value is 0.0009 which is less that 0.05. In the chi-squared test the P-Value = 0.0012 which is less than 0.05. This means death is not independent on gender. The data collected reflects that males were most likely to die from COVID than females. 
